### 2023-04-15 definitions.tex
details

In this section, we introduce general notations and preliminary definitions.

@@

The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.

%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

\mathieu{On peut probablement enlever la Proposition $8$ alors. La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
\fi

Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

@@

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

@@

For a WSTS $\mathscr{S}$ and an upward closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We then have:  $\pred^*(U) = U_{k_0}$. \\

@@

\begin{definition}{\cite{BFM-ic17}} 
A class $C$ of WSTS is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$ of an ordered transition system
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}
\end{definition}

\mathieu{I'm using the notion of ideally effective-ness for WBTS here}

\noindent
{\bf Vector addition system with states}

\begin{definition}
A {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control states} of $V$, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control transitions} of $V$. 
% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e. $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.
\end{definition}

Subsetquently, $Q \times \N^d$ denotes the set of configurations of $V$.
For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$

\problemx{$d$-VASS reachability}
{A $d$-VASS $V=(Q,T)$, and configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
{$s \to^* s'$ ? \\}

\problemx{$d$-VASS zero reachability}
{VASS $V=(Q,T)$, $q(\textbf{v}) \in Q \times \N^d$}
{$\exists p \in Q ~ q(\textbf{v}) \to^* p(\textbf{0})$? \\}

\noindent
{\bf Computing the maximal $X$-live system}

%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
\end{proof}
%
Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman.... Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman. problèmes: marquage T-bloqué, dead, T-continual ?.

>>>>

%   In this section, we introduce general notations and preliminary definitions.

@@

The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

\mathieu{On peut probablement enlever la Proposition $8$ alors. La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
\fi
%
Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

@@

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

@@

For a WSTS $\mathscr{S}$ and an upward-closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We then have:  $\pred^*(U) = U_{k_0}$. \\
%

\noindent
\subsection{Computing the maximal $X$-live system}

%		l'état atteint $s'$ est au dessus des minimaux de $\pred^*(X)$, ainsi par (upward) compatibility on sera toujours capable d'aller dans $X$. 
\end{proof}
%
%
%\noindent
%{\bf Vector addition system with states}

Let us recall that a {\em vector addition system with states (VASS)} in dimension $d$ ($d$-VASS for short) is a finite $\mathds{Z}^d$-labeled directed graph $V = (Q,T)$, where $Q$ is the set of {\em control-states}, and where $T \subseteq Q \times \mathds{Z}^d \times Q$ is the set of {\em control-transitions}. 
% The {\em size} of $V$ is defined as $|V|=|Q|+|T|*d*|log(||T||)$ where $||T||$ denotes the absolue value of the largest number that appears in $T$, i.e. $||T|| = max\{ ||\textbf{z}||: (p,\textbf{z},q) \in T\}$.

Subsetquently, $Q \times \N^d$ is the set of configurations of $V$.
For every configuration $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t = (p, \textbf{z}, q)$ we write $p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ whenever $\textbf{v} = \textbf{u} + \textbf{z} \geq \textbf{0}$

Pour les VASS, je crois qu'on a une borne (2-exp) à la taille de $\pred^*(X)$ et à la taille de ses éléments \cite{DBLP:conf/rp/BozzelliG11} et c'est directement implémentable, donc en particulier, pour tout VASS, il existe un autre VASS de taille 2-exp plus grand qui a le plus grand sous-comportement restant dans $X$ .
Je crois que ça résoud aussi un pb résolu dans Valk et Jantzen 1985 \cite{DBLP:journals/acta/ValkJ85} mais en utilisant le graphe de karp et Miller donc avec complexité ackerman.... Dans ce cas, ça permettrait de résoudre autrement les premiers des 4 pbs dans \cite{DBLP:journals/acta/ValkJ85} avec une complexité 2-exp plutôt qu'ackerman. problèmes: marquage T-bloqué, dead, T-continual ?.

@@

\alain{pas ici}
\problemx{$d$-VASS reachability}
{A $d$-VASS $V=(Q,T)$, and two configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$}
{$p(\textbf{u}) \to^* q(\textbf{v})$? \\}

\problemx{$d$-VASS zero-reachability}
{VASS $V=(Q,T)$, $p(\textbf{u}) \in Q \times \N^d$}
{$\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$? \\}

The two previous problems are well-known decidable.

### 2023-04-15 reachable-resilience.tex
details

\section{State resilience}

@@

\begin{definition}
A {\em Reset VASS} in dimension $d$ ($d$-dim Reset VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

Again $Q \times \N^d$
 denotes the set of configurations of $V$.
For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control transition $t$ we write
$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when

@@

\begin{samepage}
\problemx{$d$-dim Reset VASS zero reachability}
{A $d$-dim Reset VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
\end{samepage}

We reduce this problem to the problem of deciding whether a control transition is live, that is, whether it is always ,
which we then reduce to the state-resilience problem.
A control transition $t$ of a $d$-dim Reset VASS is called {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole $d$-dim Reset VASS is live if all its control transitions are
live. This leads to the following problem.

\problemx{$d$-dim Reset VASS $t$-liveness}
{$d$-dim Reset VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
{Is $t$ live in $s_0$ ? \\}

@@

\begin{proposition}\label{reductions}
% In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
{\sc $d$-dim Reset VASS $t$-liveness} is reducible to {\sc state-resilience}.
\end{proposition}

>>>>

\section{State-resilience}

@@

\begin{definition}
A {\em Reset VASS} in dimension $d$ (Reset $d$-VASS for short) is a finite 
% $\mathds{Z}^d$
labeled directed graph $V = (Q,T)$, where $Q$ will be referred to as the {\em control-states} of $V$, and where 
%$T \subseteq Q \times \mathds{Z}^d \times Q$
$T \subseteq Q \times Op \times Q$
 will be referred to as the {\em control-transitions} of $V$,
where $Op = \{ add(\textbf{z}) \mid \textbf{z} \in \mathds{Z}^d\} \cup 
		\{ reset(i) \mid i \in \{1,\ldots,d\} \}$.
\end{definition}

Again $Q \times \N^d$
 denotes the set of configurations of $V$.
For every configurations $p(\textbf{u}), q(\textbf{v}) \in Q \times \N^d$ and every control-transition $t$ we write
$p(\textbf{u}) \xrightarrow{t} q(\textbf{v})$ when

@@

\alain{ce qui suit est inutile}

\begin{samepage}
\problemx{Zero-reachability}
{A Reset $d$-VASS $V=(Q,T)$, $q(\textbf{u}) \in Q \times \N^d$}
{$\exists p \in Q ~ q(\textbf{u}) \to^* p(\textbf{0})$? \\}
\end{samepage}
%
\alain{ce qui précède est inutile}
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live, that is, whether it is always ,
which we then reduce to the state-resilience problem.
A control-transition $t$ of a Reset $d$-VASS is {\em live} in a configuration $s$ if for each $q(\textbf{v}) \in \post^*(s)$ there exists a \alain{mélange de s et (p,u)}
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole Reset $d$-VASS is live if all its control-transitions are
live. This leads to the following problem.

\problemx{$t$-liveness}
{Reset $d$-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
{Is $t$ live in $s_0$ ? \\}

@@

\begin{proposition}\label{reductions}
% In WSTS, $t$-liveness is reducible to {\sc state-resilience}.
{\sc $t$-liveness} is reducible to {\sc state-resilience} in Reset $d$-VASS.
\end{proposition}

### 2023-04-15 resilience-safe-up.tex
details

Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$. The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo. $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set). Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.

>>>>

\alain{pas ici}
\begin{definition}{\cite{BFM-ic17}} 
A class $C$ of WSTS is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$ of a WSTS
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}
\end{definition}

\mathieu{I'm using the notion of ideally effective-ness for WBTS here}
%
Let us recall that the \emph{completion}  \cite{BFM-ic17} of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ is the associated ordered transition system $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ where states of $\hat{\mathscr{S}}$ are ideals of $S$ and $I \rightarrow J$ if $J$ belongs to the finite ideal decomposition of $\downarrow \post_{\mathscr{S}}(I)$. The completion is always finitely branching but it is not necessarly WSTS since $\subseteq$ is not necessarly a wqo. $\hat{\mathscr{S}}$ is WSTS iff $\mathscr{S}=(S,\rightarrow, \leq)$ is $\omega^2$-WSTS (intuitively speaking, $(S,\leq)$ must not contain the Rado set). Coverability is shown decidable  [Theorem 44] in \cite{BFM-ic17} for completion-post-effective $\omega^2$-WSTS.

### 2023-04-15 resilience.tex
details

Surprinsingly, the decidability of resilience problems was not studied in the previous papers.

@@

\begin{example}
% Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
Let us consider now a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
then resilience hold (one always ends up in the positive integers) but not bounded resilience
(for every bound $b$ there is an element of $\Bad$ e.g. $q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$). 
\end{example}

>>>>

Surprinsingly, we have not found decidability results about the resilience problems for WSTS.

@@

\begin{example}
% Si on prends un Z-vass avec même seulement un compteur, Safe les entiers positifs, Bad le reste, avec juste une transition qui fait +1, alors on a la résilience (on reviens toujours dans les positifs) mais pas la bounded resilience.
Let us consider now a $\mathds{Z}$-VASS with only one state $q$, one counter, and one transition that adds $+1$ to the counter, with $\Safe = \uparrow q(0)$, $\Bad = \downarrow q(-1)$,
then resilience hold (one always ends up in the positive integers) but not bounded resilience: for every bound $b$ there is an element of $\Bad$ e.g. $q(-b-1)$ which necessitate at least $b+1$ steps before it can reach $\Safe$. 
\end{example}
